Classification of daily crop phenology in PhenoCams using deep learning and hidden markov models.

### 2022-01-06 manuscript/main.tex
2nd revision
Accepted
only a minor edit for model comparison

We used the VGG16 model in the Keras python package to classify images into the 21 classes (Table~\ref{table1})\citep{simonyan2014-VGG16, chollet2018-keras}. The model allows us to specify the hierarchical structure of the three categories, such that the predicted class probabilities for any image sum to one within each category. We held out 20\% of the images as a validation set. The validation set included all images from three cameras: arsmorris2, mandani2, and cafboydnorthltar01 totaling 10,172 images. It also included 8,124 randomly selected images from the remaining locations to obtain the full 20\%. This resulted in a validation sample size of 18,296 and a training sample size of 73,188. The VGG16 classification model was trained fully, as opposed to using transfer learning \citep{norouzzadeh2018}. We experimented with transfer learning, where a pre-trained model is fine-tuned using our own data, but found that training the model fully had better results.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\supplementary{The following are available online at \linksupplementary{s1}, Figure S1:  Locations of all PhenoCam sites used in the study, Figures S2-S3: Confusion matrices for the classification results, Figures S4-S13: Example classification output from the VGG16 classifier, Table S1: Site information for PhenoCams used in this study, Tables S2-S3: Transition matrices for the hidden markove models.}

>>>>

We used the VGG16 model in the Keras python package to classify images into the 21 classes (Table~\ref{table1})\citep{simonyan2014-VGG16, chollet2018-keras}. We evaluated several other models preconfigured in the Keras package and found VGG16 to be the best performing (see Supplement). The model allows us to specify the hierarchical structure of the three categories, such that the predicted class probabilities for any image sum to one within each category. We held out 20\% of the images as a validation set. The validation set included all images from three cameras: arsmorris2, mandani2, and cafboydnorthltar01 totaling 10,172 images. It also included 8,124 randomly selected images from the remaining locations to obtain the full 20\%. This resulted in a validation sample size of 18,296 and a training sample size of 73,188. The VGG16 classification model was trained fully, as opposed to using transfer learning \citep{norouzzadeh2018}. We experimented with transfer learning, where a pre-trained model is fine-tuned using our own data, but found that training the model fully had better results.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\supplementary{The following are available online at \linksupplementary{s1}, Figure S1:  Locations of all PhenoCam sites used in the study, Figures S2-S3: Confusion matrices for the classification results, Figures S4-S13: Example classification output from the VGG16 classifier, Figures S14-S15: CNN Model comparison figures, Table S1: Site information for PhenoCams used in this study, Tables S2-S3: Transition matrices for the hidden markove models.}
